Robust Clustering for Ad Hoc Cognitive Radio Network

### 2017-09-18 journal-draft/response_to_editors.tex

\section{Reviewer 1:}

\begin{enumerate}

\item \textcolor{blue}{  First of all, clustering technique is not new and it is one of the well-researched areas in ad hoc network, cognitive radio, HetNet etc. The article does not review the latest publications in this area. The references are not adequate due to the fact that no recent works have been investigated. Therefore, it makes it insufficient (or unsatisfactory) work in present context.}

\textbf{Answer:} As mentioned in the introduction section, there has been a lot of works on clustering in ad hoc networks or sensor networks.
They address problems such like routing, power consumption and communication quality.
Compared with the wireless networks where spectrum availability is not opportunistic, the clustering in CRN facilitated the better usage of the available spectrum, meanwhile clustering itself is restricted by the spectrum availability.
This fact makes clustering in CRN a difficult topic.
Up to now, only \cite{LIU_TMC11_2, Li11_ROSS, mansoor_15_cluster_robust} have tried to address the robustness issue of the formed clusters in CRN.
After searching, we find one relavent paper \cite{Mansoor2015} from 2015 IEEE Crowncom.
It proposes a clustering algorithm which aims to speed up the process of re-clustering in case a cluster doesn't uphold in from of active primary users.
But this work doesn't consider to improve the ability of the formed clusters to maintain, besides, the messaging cost which is involved in this process is not considered.
We have included this citation in the related work section in the revised version.

@@

\item \textcolor{blue}{  The presentation of the paper is not good enough and needs a lot of improvements. For instance, ROSS scheme has been mentioned in the paper in many places, but it is not properly defined (though reference has been given). I had to check the reference paper to know it is Robust Spectrum Sharing. The journal article must be self-contained.}

    \textbf{Answer:} The explanation of the scheme name is provided in the introduction section.
    The paper has been thoroughly proofread.

@@

\item \textcolor{blue}{  The paragraphs are too long in the Introduction. As a result, the paragraph contains too many information which are not related each other. This makes readers difficult to understand the concept.}

    \textbf{Answer:} We polish the paper and now the section of introduction consists of 4 paragraphs: introduction of CRN with mentioning of spectrum sensing $\rightarrow$ clustering in CRN $\rightarrow$ robustness of clustering $\rightarrow$ the role of cluster size in robust clustering in CRN.
    The part of related works becomes one separate section.

@@

\item \textcolor{blue}{   In page 3, it is mentioned that “A cluster can be formed only by the cluster head”. As a matter of fact, the cluster head formation is itself a very challenging task in any cluster based approach. There are many previous works which describe the cluster head selection techniques. Please justify your statement with proper assumption or previous works that how the cluster heads are selected. I can see the description is in the few pages later but some information must be provided where it is described first time. By the way, the CH formation technique described in 4.1.1 is not novel.}

\textbf{Answer:} We agree with the reviewer that this statement is not well explained.
	    At first, we put this line in the first submission and wanted to say that after the clusters are decided, it is the cluster head's job to form a cluster out of the neighboring nodes.
In the revised version, we have changed the statement as "Our scheme enables the cluster heads are selected in a distributed manner, and a cluster can only be formed only by these selected cluster head. "

@@

\item \textcolor{blue}{  In my understanding, the values of connectivity degree in Fig. 1 came from the assumption of $K_A$, $K_B$ …..$K_H$. Each node has number of available channels and I assume that they have been randomly chosen as mentioned in Fig. 1 Caption. It would be good to mention how to obtain such information in ad hoc cognitive radio network?}

\textbf{Answer:} 
The available channels on the nodes in Fig. 1 is randomly chosen, which is only used to explain our scheme.
But in the simulation, the available channels on the secondary users are decided by the primary users.
As stated in the system model, both primary and second users work on the same set of channels, i.e., the former use the channel whenever it is needed, and the later can only use it opportunistically.
    We assume the primary users are randomly located and work on each channel with a certain probability.
    With spectrum sensing, the secondary users can detect the presence of active primary users, and thus obtain the available channels.

@@

\item \textcolor{blue}{  In page 7, what does “the toy network in Figure 2” mean? I hope it’s a typo. Otherwise, the Sections 4.2.1, 4.2.2, 4.2.3 are nicely written.}

\textbf{Answer:} It is a typo and we have changed the sentence to "Now we apply both ROSS-DGA and ROSS-DFA to the network in Figure 2 which has been applied the phase I of ROSS."

\item \textcolor{blue}{  In simulation, have authors considered the spectrum sensing parameters on each secondary nodes? Are the channels randomly made available to users? Please clarify. What value of A is chosen in simulation?}

\textbf{Answer:} In the most part of the simulation, we assume the secondary users have accurate spectrum sensing ability.
They are able to detect the active primary users in their vicinity.
We can say the channel availability on the secondary users is decided by the primary users.
As to a secondary user, if none of the primary users whose transmission range include the secondary user works on a channel, that channel will be available for that secondary user. 
We also consider the existence of sensing error, in Section 6.2.4 in new version, we investigate the influence caused by false negative sensing error.

 'A' represents the length the simulation square.
    And the radius of transmission of both primary and secondary users is a fraction of A.
    By doing this, we separate our scheme with concrete physical layer communication technologies, and A can be given a concrete value in practice when the physical layer technology is decided.

@@

\item \textcolor{blue}{ It is mentioned in the article that “The number of licensed channels is 10, each PU is operating on each channel with probability of 50\%”. It means there will be only 5 channels available (maximum) to each secondary users. In section 5.1, 7 channels are available for each CR node. It means there is chance of interference in two nodes. Please clarify those conflicts.}

\textbf{Answer:} 
The simulation consists of two parts, the comparison between the centralized and distributed schemes, and the comparison between the distributed schemes.
Some parameters are identical for both parts (stated in the second paragraph in Section 6), but there are some parameters are different (as stated in the beginning of Section 6.1 and 6.2 respectively).
The difference lies in the number of the CRN and the primary users, and the transmission ranges.
The reason is because there doesn't exist an efficient method to solve the centralized scheme, thus we have to do simulation in a smaller network.

The parameter setting questioned by the reviewer is for the simulation involving centralized schemes in Section 6.1.
%According to the description, here will be "only 5 channels available", not "to each secondary users", but to the nodes within the primary user's transmission range.
    Please note that the transmission radius of primary user is A/3, which means some nodes are not affected by all the primary users.
    Other than that, from the simulation we know averagely 7.1 channels are available for each secondary user.

@@

\section{Reviewer 2:}

Comments to the Author
In this paper, authors investigate the optimization algorithms for node clustering strategy in ad hoc based cognitive radio networks, where both the centralized solution and distributed solution are considered.
\begin{enumerate}

\item \textcolor{blue}{ In Section 2, that the definition of channel $k\in K$, i.e. fading channel or noisy channel? Does the channel contain pathloss and shadowing and/or fast fading? Does the value of k indicate the channel gain in either time or frequency domain? }

\textbf{Answer:} As shown in the section of system model, "We adopt the unit disk model for both primary and CR users’ transmission."
In other words, we don't make any assumption on the concrete propagation model, but just assume the communication is successful when the two ends locate within each other's transmission range.
We adopt this simple communication model so as to focus on secondary users' behavior to pursue the robust cluster structure.

@@

\item \textcolor{blue}{ Does the channel invariant or variant (how fast) during the clustering for how long windows range? In practical, the channel changes along with the transmission, the performance clustering may be impacted by the varying channel. What's the impact of insufficient channel knowledge?}

\textbf{Answer:} We assume the channel availability doesn't change during the clustering process.
We agree with the reviewer that the channel variance affects the clustering by disrupting the communication.
In reality, if the communication during the clustering process doesn't succeed as the channel quality deteriorates, the involved nodes can still make decisions based on the available knowledge of neighbors according to our algorithm.

\item \textcolor{blue}{ The work emphasize on the robustness of clustering. However, such an evaluation in terms of robustness metric is not clear. The definition of robustness and relative discussions as well as the results should be further elaborated. The definition appears in Section 5 but my suggestion is to clearly state it in the system model and describe why choose this robustness metric according its relation with system design.}

\textbf{Answer:} We take the advice and have done the modifications in two places.
1. We mention the definition of robustness of clusters in the 3rd paragraph of the introduction section.
2. In Section 3.1, the robustness is associated with the number of secondary users which are not included in any cluster.
with these modifications, readers will understand the robustness metric in Section 6 (Section 5 in the previous version) easily.

\item \textcolor{blue}{ In Section 4, algorithm flow charts should be included.}

\textbf{Answer:} a flow chart as Figure 1 is added in Section 5 (which is Section 4 before this major revision)

\item \textcolor{blue}{ In Section 5, Fig. 10 is missing. Fig. 14-16 can be subfigures, similar to Fig. 17-19.}

\textbf{Answer:} Advice is taken and change is made. Please refer Fig 8 - 9.

\end{enumerate}

@@

\section{Reviewer 3:}

Comments to the Author
In this paper, the authors consider the robust clustering problem in the ad-hoc Cognitive Radio Networks (CRNs). The paper is well formalized. The research background and motivation are well described. The research problem is well addressed. The simulation based experiment is conducted to show the performance of the suggested solution approach to the addressed problem.

Overall, the considered topic is very interesting. The authors well presented the considerations and assumptions on several key research issues, which inherently exist in the ad-hoc networks, for instance, decentralized spectrum sensing, synchronization on the sensing results and congestion control. To solve the addressed problem, the authors jointly use different research methods such as graphic theory and game theory, which have been widely reported in recent studies on CRNs. The experiment is well conducted with the thorough performance analysis and comparison on different experimental scenarios.
\begin{enumerate}
\item \textcolor{blue}{ However, it seems that the authors totally miss a very important research issue, which is referred to as spectrum sensing errors in terms of overlook and misidentification on the channel availability by Secondary Users (SUs). Because the sensing errors may highly affect the effectiveness and performance of the suggested solution approach, the authors need to complement the paper by accordingly either advancing the feasible solutions to alleviate sensing errors or analyzing the effect of sensing errors on the suggested solution approach.}

 \textbf{Answer:} It is true that the sensing error was not considered in the first submission.
We agree with the reviewer that this is an important issue, as sensing error is inevitable in practical applications.
We have done additional simulations about the effects of sensing errors on the distributed schemes in Section 6.2.4.
\end{enumerate}

### 2017-09-23 journal-draft/response_to_editors.tex

\section{Reviewer 1:}

\begin{enumerate}

\item \textcolor{blue}{  First of all, clustering technique is not new and it is one of the well-researched areas in ad hoc network, cognitive radio, HetNet etc. The article does not review the latest publications in this area. The references are not adequate due to the fact that no recent works have been investigated. Therefore, it makes it insufficient (or unsatisfactory) work in present context.}

\textbf{Answer:} As mentioned in the introduction section, there has been a lot of works on clustering in ad hoc networks or sensor networks.
They address problems such as routing, power consumption and communication quality.
Compared with the wireless networks where spectrum availability is not opportunistic, the clustering in CRN facilitated the better usage of the available spectrum, meanwhile clustering itself is restricted by the spectrum availability.
This fact makes clustering in CRN a difficult topic.
Up to now, only \cite{LIU_TMC11_2, Li11_ROSS, mansoor_15_cluster_robust} have tried to address the robustness issue of the formed clusters in CRN.
After searching, we find one additional relevant paper \cite{Mansoor2015} from 2015 IEEE Crowncom.
It proposes a clustering algorithm which aims to speed up the process of re-clustering in case a cluster doesn't uphold in from of active primary users.
But this work doesn't consider to improve the ability of the formed clusters to be maintained, besides, the messaging cost which is involved in this process is not considered.
We have included this citation in the related work section in the revised version.
We will appreciate if the reviewer could provide more specific references that we have not mentioned yet.

@@

\item \textcolor{blue}{  The presentation of the paper is not good enough and needs a lot of improvements. For instance, ROSS scheme has been mentioned in the paper in many places, but it is not properly defined (though reference has been given). I had to check the reference paper to know it is Robust Spectrum Sharing. The journal article must be self-contained.}

    \textbf{Answer:} In the revised version, we explain the name ROSS in the fifth paragraph when we introduce our scheme.
%    The paper has been thoroughly proofread.

### 2017-09-23 journal-draft/ettdoc.tex
moving some text from section 2 to 1

Cluster size is inherently correlated with the cluster robustness.
Given a CRN, it is obvious that getting rid of certain secondary users from a cluster may yield more CCs.
However, this contradicts to one of the motivations of clustering, \ie the cooperative decision making, where more users result in more accurate sensing result~\cite{Consensus_based_clustering12}.
Thus the number of small clusters, especially the \textit{singleton clusters}, \ie the cluster which has only one CR node, should be minimized.
On the other hand, as to the cluster which consists of many nodes, although the spectrum sensing is benefited, it usually leads to less common channels. This undermines the robustness of the cluster.
Thus, the cluster robustness discussed in terms of number of CCs carries little meaning when the sizes of formed clusters are not considered.
Cluster sizes also play a role in transmission power consumption, \ie cluster size affects the transmission power consumption when routing is conducted~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
In this paper we will analyze the robust clustering problem and propose efficient solutions.
Our schemes generate the clusters which demonstrate both the robustness against the increasing primary activity and the compliance with the desired cluster sizes.
Hence our schemes provide the network designers with more freedom when the power consumption is involved in the design of network applications.
%In this paper we will introduce a robust cluster structure for CRN, where the robustness means the ability of the clusters to sustain under the influence of the increasing activity of primary users, and the direct metric is the number of CCs in the clusters.

\section{Related Works and the Relation with This Paper}
In the following we will give a brief review on the clustering in CRN, with emphasis on the theme of robust clustering.
%%[crn for clustering] 
%%due to attenuation of signal propagation, primary users can only be detected by CR users when they locate closely to CR users.
%In cognitive radio networks, secondary users which locate closely with each other are possibly affected by the same group of primary users, so that the availability of licensed spectrum is similar to them, \ie certain channels are available on each of them.
%The similarity of available spectrum on a group of neighbouring CR nodes, along with the benefit of collaborative decision among multiple nodes, leads to clustering as an effective approach for many applications.
%[robustness issue for clustering]
%%%%%%%%%%%
%On the other hand, large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members in a large cluster is a substantial burden. 
%Hence, the cluster size should fall in a desired range according to different application scenarios~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
%%%%%%%%%%%%
%with the intention to provide a robust infrastructure for the applications which may be conducted above that in CRN.
%that being said, when we design this clustering scheme in CRN, we do consider the services to be delivered by the network.
%Cluster sizes affect the power consumption directly in wirelss sensor networks~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
%When the cluster size is not deliberately discussed, it is also decided alongside the optimization of the network, \eg the opimal selection of the aggregation points in a wireless sensor network results in the clusters~\cite{data_aggr_algorithm_04}.
%%%%%%%%%%%%%
%Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
With regard to forming clusters in CRN, deciding on the common channel within each cluster is the foremost question to answer.
\cite{Zhao07, Chen07,Affinity_clustering_09icccn} propose the clustering schemes and enforce that every cluster possesses at least one CC.
Clustering scheme~\cite{Consensus_based_clustering12} looks for a network partition which improves the accuracy of spectrum sensing with the cluster structure.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{TWC2012_cooperative_communication} forms the clusters by deciding on the cluster heads, where the transmit power for the long-haul transmission between the cluster heads is minimized.
\cite{clustering_globecom11} proposes a cluster structure which promises energy efficiency.
\cite{cluster_EW10} proposes strategy on how to decide on the CCs and access the multiple CCs within clusters.
An event-driven clustering scheme is proposed for cognitive radio sensor networks in \cite{Ozger_cluster_crsn_13}.
None of the above mentioned schemes provides a certain robustness to formed clusters against primary users.

The authors of~\cite{Mansoor2015} propose a clustering algorithm which aims to speed up the process of re-clustering in case a cluster doesn't uphold in from of active primary users.
But this work doesn't consider to improve the ability of the formed clusters to maintain, and the messaging cost which is involved in this process is not considered.
Robustness of clusters is discussed in~\cite{LIU_TMC11_2}. A distributed clustering scheme (denoted as SOC) which is designed to generate robust clusters against primary users is proposed.
The robustness of the clusters comes from the fact that there are multiple CCs available for the clusters.
SOC involves three phases of distributed executions.
In the first phase, every secondary user forms clusters with some one-hop neighbors. In the second and third phase, each secondary user seeks to either merge other clusters or join one of them.
The metric adopted by every secondary user in all the phases is the product of the number of CCs and cluster size.
The drawbacks of this scheme are as follows, although the adopted metric considers both the cluster size and the number of CCs, cluster formation can be easily dominated by only one factor.
For example, a node which accesses abundant channels may form a cluster by itself and doesn't have motivation to unite neighboring nodes to form a cluster.
In addition, this scheme leads to the high variance of the cluster sizes, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.
\cite{mansoor_15_cluster_robust} presents a heuristic method to form clusters. Although the authors claim that robustness is one goal to achieve, the minimum number of clusters is finally pursued.
A distributed clustering scheme ROSS (Robust Spectrum Sharing) is proposed in~\cite{Li11_ROSS} under the game theoretic framework. 
Compared with the clustering schemes introduced above, the clusters are formed faster and the clusters possess more CCs than SOC.
However, as all the other clustering schemes, this scheme does not have control over the formation of the very small or very large clusters which are not desirable.
Furthermore, both this work and SOC define robustness just to be the number of CCs per cluster, and don't consider the sustainability of clusters against the increasing activity of primary users, which leaves their claim on robustness unverified.
This paper fundamentally extends~\cite{Li11_ROSS} in two dimensions.
First, this paper considers the correlation between the cluster robustness and the cluster size.
As a result the cluster robustness is no longer indicated merely by the average number of CCs of clusters, but the ability of the CR nodes to get benefit from collective decision and of the clusters to sustain when the primary users' activity becomes more dynamic.
Besides, this involment of cluster size enable us to solve the problem of clusters size devergence, which appears in~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.
Second, this paper provides a comprehensive analysis of the robust clustering problem, proposes both centralized and distributed solutions, and conducts extensive simulations. 
%We stick to the motivation of forming robust clusters in CRN \ie let more CR users benefited from the cooperative decision making which is due to the clusters. 
%We propose both centralized and distributed clustering schemes, which result in more CR users in the clusters composed with multiple CR users, not only when the clusters are just formed, but also afterwards when the primary users' activity changes.
%Besides, both centralized and distributed schemes take the cluster size into consideration, and both can control the sizes of the formed clusters to certain extent.
%In particular, the decentralized clustering approach ROSS (RObust Spectrum Sharing) is able to form clusters with desired sizes.
%When compared with other distributed works, ROSS involves smaller signaling overhead and the generated clusters are significantly more robust against the primary users which appear after the clusters are formed.
%thesis
%and the generated clusters are more robust than the other clustering scheme which has claims on cluster robustness, \ie more secondary users residing in clusters against increasing influence from primary users.
%ROSS selects cluster head through coordination within its neighborhood, and then cluster membership is decided locally and its fast convergence is proved under game theoretic framework. 
%For our scheme we can prove convergence in cluster formation phase and resolve ambiguities with respect to cluster membership in a game-theoretic setting. 
The new extensions are made on basis of ROSS and its light weight version, the latter involves less overheads thus is more suitable for the scenario where fast deployment is desired.
Throughout this paper, we refer to the clustering schemes on the basis of ROSS as the \textit{variants of ROSS}.

The rest of paper is organized as follows. 
We present the system model and the robust clustering problem in Section~\ref{sec:model}. 
The centralized and distributed solutions are introduced in Section~\ref{centralized_solution} and~\ref{ross} respectively.
%The clustering problem is given through analysis and a centralized scheme is proposed in section~\ref{centralized_scheme}.
Some Extensive performance evaluation is presented in Section~\ref{performance}.
Finally, we conclude our work and point out the direction for future research in Section~\ref{conclusion}.

>>>>

Cluster size is inherently correlated with the cluster robustness.
Given a CRN, it is obvious that getting rid of certain secondary users from a cluster may yield more CCs.
However, this contradicts to one of the motivations of clustering, \ie the cooperative decision making, where more users result in more accurate sensing result~\cite{Consensus_based_clustering12}.
Thus the number of small clusters, especially the \textit{singleton clusters}, \ie the cluster which has only one CR node, should be minimized.
On the other hand, as to the cluster which consists of many nodes, although the spectrum sensing is benefited, it usually leads to less common channels. This undermines the robustness of the cluster.
Thus, the cluster robustness discussed in terms of number of CCs carries little meaning when the sizes of formed clusters are not considered.
Cluster sizes also play a role in transmission power consumption, \ie cluster size affects the transmission power consumption when routing is conducted~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.

In this paper robust clustering in CRN is analyzed and both centralized and distributed schemes are proposed.
This paper is based on our previous work ROSS (Robust Spectrum Sharing)~\cite{Li11_ROSS}, but fundamental extensions are made in several aspects.
This paper provides a comprehensive analysis of the robust clustering problem, analyze and investigate the influence of cluster size and the number of CCs on cluster robustness.
We claim that the cluster robustness against the primary users is no longer indicated merely by the average number of CCs of clusters, but by the ability of the clusters to uphold when the primary users' activity becomes more active.
The involment of cluster size enables us to solve the problem of clusters size devergence, which appears in~\cite{Li11_ROSS} and~\cite{LIU_TMC11_2}.
These new extensions are made on the basis of ROSS and its light weight version, the latter involves less overheads thus is more suitable for the scenario where fast deployment is desired.
Throughout this paper, we refer to the clustering schemes on the basis of ROSS as the \textit{variants of ROSS}.

%Our schemes generate the clusters which demonstrate both the robustness against the increasing primary activity and the compliance with the desired cluster sizes.

The rest of paper is organized as follows. 
In section~\ref{related_work}, we will introduce the clustering techniques in CRN and those focusing on robustness.
The relation between this paper and work~\cite{Li11_ROSS} is elaborated.
She system model and the robust clustering problem formulation are presented in Section~\ref{sec:model}. 
The centralized and distributed solutions are introduced in Section~\ref{centralized_solution} and~\ref{ross} respectively.
%The clustering problem is given through analysis and a centralized scheme is proposed in section~\ref{centralized_scheme}.
Extensive performance evaluation is given in Section~\ref{performance}.
Our work is concluded in Section~\ref{conclusion}.

\section{Related Works and the Relation with This Paper}
\label{related_work}
In the following we will give a brief review on the clustering in CRN, with emphasis on the theme of robust clustering.
%%[crn for clustering] 
%%due to attenuation of signal propagation, primary users can only be detected by CR users when they locate closely to CR users.
%In cognitive radio networks, secondary users which locate closely with each other are possibly affected by the same group of primary users, so that the availability of licensed spectrum is similar to them, \ie certain channels are available on each of them.
%The similarity of available spectrum on a group of neighbouring CR nodes, along with the benefit of collaborative decision among multiple nodes, leads to clustering as an effective approach for many applications.
%[robustness issue for clustering]
%%%%%%%%%%%
%On the other hand, large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members in a large cluster is a substantial burden. 
%Hence, the cluster size should fall in a desired range according to different application scenarios~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
%%%%%%%%%%%%
%with the intention to provide a robust infrastructure for the applications which may be conducted above that in CRN.
%that being said, when we design this clustering scheme in CRN, we do consider the services to be delivered by the network.
%Cluster sizes affect the power consumption directly in wirelss sensor networks~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
%When the cluster size is not deliberately discussed, it is also decided alongside the optimization of the network, \eg the opimal selection of the aggregation points in a wireless sensor network results in the clusters~\cite{data_aggr_algorithm_04}.
%%%%%%%%%%%%%
%Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
With regard to forming clusters in CRN, deciding on the common channel within each cluster is the foremost question to answer.
\cite{Zhao07, Chen07,Affinity_clustering_09icccn} propose the clustering schemes and enforce that every cluster possesses at least one CC.
Clustering scheme~\cite{Consensus_based_clustering12} looks for a network partition which improves the accuracy of spectrum sensing with the cluster structure.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{TWC2012_cooperative_communication} forms the clusters by deciding on the cluster heads, where the transmit power for the long-haul transmission between the cluster heads is minimized.
\cite{clustering_globecom11} proposes a cluster structure which promises energy efficiency.
\cite{cluster_EW10} proposes strategy on how to decide on the CCs and access the multiple CCs within clusters.
An event-driven clustering scheme is proposed for cognitive radio sensor networks in \cite{Ozger_cluster_crsn_13}.
None of the above mentioned schemes provides a certain robustness to formed clusters against primary users.

The authors of~\cite{Mansoor2015} propose a clustering algorithm which aims to speed up the process of re-clustering in case a cluster doesn't uphold in from of active primary users.
But this work doesn't consider to improve the ability of the formed clusters to maintain, and the messaging cost which is involved in this process is not considered.
Robustness of clusters is discussed in~\cite{LIU_TMC11_2}. A distributed clustering scheme (denoted as SOC) which is designed to generate robust clusters against primary users is proposed.
The robustness of the clusters comes from the fact that there are multiple CCs available for the clusters.
SOC involves three phases of distributed executions.
In the first phase, every secondary user forms clusters with some one-hop neighbors. In the second and third phase, each secondary user seeks to either merge other clusters or join one of them.
The metric adopted by every secondary user in all the phases is the product of the number of CCs and cluster size.
The drawbacks of this scheme are as follows, although the adopted metric considers both the cluster size and the number of CCs, cluster formation can be easily dominated by only one factor.
For example, a node which accesses abundant channels may form a cluster by itself and doesn't have motivation to unite neighboring nodes to form a cluster.
In addition, this scheme leads to the high variance of the cluster sizes, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.
\cite{mansoor_15_cluster_robust} presents a heuristic method to form clusters. Although the authors claim that robustness is one goal to achieve, the minimum number of clusters is finally pursued.
%
A distributed clustering scheme ROSS (Robust Spectrum Sharing) is proposed in~\cite{Li11_ROSS} under the game theoretic framework. 
Compared with the clustering schemes introduced above, the clusters are formed faster and the clusters possess more CCs than SOC.
However, as all the other clustering schemes, this scheme does not have control over the formation of the very small or very large clusters which are not desirable.
Furthermore, both this work and SOC define robustness just to be the number of CCs per cluster, and don't consider the sustainability of clusters against the increasing activity of primary users, which leaves their claim on robustness unverified.
%
